---
author: "Piotr Patrzylas"
title: "Bike Sharing - an example of linear regression"
output: html_document
---

"Readme.txt" file found on the: http://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset

=========================================
Background 
=========================================  

Bike sharing systems are new generation of traditional bike rentals where whole process from membership, rental and return 
back has become automatic. Through these systems, user is able to easily rent a bike from a particular position and return 
back at another position. Currently, there are about over 500 bike-sharing programs around the world which is composed of 
over 500 thousands bicycles. Today, there exists great interest in these systems due to their important role in traffic, 
environmental and health issues. 

Apart from interesting real world applications of bike sharing systems, the characteristics of data being generated by
these systems make them attractive for the research. Opposed to other transport services such as bus or subway, the duration
of travel, departure and arrival position is explicitly recorded in these systems. This feature turns bike sharing system into
a virtual sensor network that can be used for sensing mobility in the city. Hence, it is expected that most of important
events in the city could be detected via monitoring these data.

=========================================
Data Set
=========================================  
Bike-sharing rental process is highly correlated to the environmental and seasonal settings. For instance, weather conditions,
precipitation, day of week, season, hour of the day, etc. can affect the rental behaviors. The core data set is related to  
the two-year historical log corresponding to years 2011 and 2012 from Capital Bikeshare system, Washington D.C., USA which is 
publicly available in http://capitalbikeshare.com/system-data. We aggregated the data on two hourly and daily basis and then 
extracted and added the corresponding weather and seasonal information. Weather information are extracted from http://www.freemeteo.com. 

=========================================
Associated tasks
=========================================  

	- Regression: 
		Predication of bike rental count hourly or daily based on the environmental and seasonal settings.  
	
	
End of "readme.txt".
	
=========================================
1. Data analysis and visualization
=========================================  
a) Daily data set:  
- instant: record index  
- dteday : date  
- season : season (1:winter, 2:spring, 3:summer, 4:fall)  
- yr : year (0: 2011, 1:2012)  
- mnth : month ( 1 to 12)  
- holiday : weather day is holiday or not (extracted from [Web Link])  
- weekday : day of the week  
- workingday : if day is neither weekend nor holiday is 1, otherwise is 0.  
+ weathersit :  
- 1: Clear, Few clouds, Partly cloudy, Partly cloudy  
- 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist  
- 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds  
- 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog  
- temp : Normalized temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-8, t_max=+39 (only in hourly scale)  
- atemp: Normalized feeling temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-16, t_max=+50 (only in hourly scale)  
- hum: Normalized humidity. The values are divided to 100 (max)  
- windspeed: Normalized wind speed. The values are divided to 67 (max)  
- casual: count of casual users  
- registered: count of registered users  
- cnt: count of total rental bikes including both casual and registered   

Data pre-processing  
```{r, eval=TRUE}
Bike1 = read.csv("day.csv")
Bike1$season <- as.factor(Bike1$season)
Bike1$dteday <- as.POSIXct(Bike1$dteday)
Bike1$yr <- as.factor(Bike1$yr)
Bike1$mnth <- as.factor(Bike1$mnth)
Bike1$holiday <- as.factor(Bike1$holiday)
Bike1$weekday <- as.factor(Bike1$weekday)
Bike1$workingday <- as.factor(Bike1$workingday)
Bike1$weathersit <- as.factor(Bike1$weathersit)
```
Data structure and summary  
```{r, echo=FALSE}
summary(Bike1)
str(Bike1)
print("NA elements")
sapply(Bike1, function (x) sum(is.na(x)))
print("Null elements")
sapply(Bike1, function (x) sum(is.null(x)))
```

Correlation between response (total count) and some of the predictors  
``` {r, echo=FALSE}
pairs(Bike1[, c(2,16)])
print("In general there is more rentals in the 2nd year. Rentals seems to follow a trend described below.")
pairs(Bike1[, c(3,5,16)])
print("Here is more clearly that for both years more bikes are hired during warmer seasons/months.")
pairs(Bike1[, c(9,10,11,16)])
print("No wonder that temperature and weather situation has positive effect on the number of bikes rented.")
pairs(Bike1[, c(12,13,16)])
print("However, humidity and windspeed have no obvious effect on the count.")


```



=========================================
2. Simple Linear Regression models
=========================================  

```{r, eval=TRUE, echo=TRUE}
plot(Bike1$cnt, Bike1$temp) # Seems to be a positive linear relationship
model <- lm(cnt ~ temp, data = Bike1)
summary(model)
# Temp p-value very low, so it is statistically siginificant.
# B0 = 1214 and B1 is 6640. This can be translated that every 
# RSE 1509 which means prediction on the response (count) based on the temperature would be off by about 1509.
# Relatively low R squared value, only 39% of the sample variation in the response is explained by the temperature.
```
  
  
=========================================
4. Multiple Linear Regression
=========================================  

```{r, eval=TRUE, echo=TRUE}
model2 <- lm(cnt ~ dteday+season+weathersit+temp+hum+windspeed, data = Bike1)
summary(model2)
```

